Functional pant indicators - wetland, indicator documentation

### 2023-04-14 DRAFT_functional plant indicators_wetland.Rmd
wetland fpi result analyses and maps

```{r}
head(results.wetland[['2-sided']])
```

>>>>

```{r}
head(results.wetland[['2-sided']])
```
#### Scaled value analyses

```{r}
#### plotting scaled values by main ecosystem type ####
## continuing with 2-sided
res.wetland <- results.wetland[['2-sided']]

# make long version of the scaled value part
res.wetland <-
  res.wetland %>% 
  pivot_longer(
    cols = c("Light1","Light2","Moist1","Moist2","pH1","pH2","Nitrogen1","Nitrogen2"),
    names_to = "fp_ind",
    values_to = "scaled_value",
    values_drop_na = TRUE
  )

# summarizing the indicator scores
res.wetland %>%
  group_by(fp_ind) %>%
  dplyr::summarize(Mean = mean(scaled_value, na.rm=TRUE))

# making the plot
ggplot(res.wetland, aes(x=factor(hovedtype_rute), y=scaled_value, fill=fp_ind)) + 
  geom_hline(yintercept=0.6, linetype="dashed") + 
  geom_violin() +
#  geom_boxplot(width=0.2, color="grey") +
  geom_point(size=0.7, shape=16, color="grey") +
  facet_wrap(~factor(fp_ind,levels=c("Light1","Moist1","pH1","Nitrogen1","Light2","Moist2","pH2","Nitrogen2")), ncol = 4) + 
  xlab("Main ecosystem type") + 
  ylab("Scaled indicator value")

#### getting scaled AND original values ####
res.wetland <- results.wetland[['2-sided']]

# make long version of the scaled value part
res.wetland <-
  res.wetland %>% 
  pivot_longer(
    cols = c("Light1","Light2","Moist1","Moist2","pH1","pH2","Nitrogen1","Nitrogen2"),
    names_to = "fp_ind",
    values_to = "scaled_value",
    values_drop_na = TRUE
  )

# add original values as well
res.wetland <- 
  res.wetland %>% add_column(original = results.wetland[['original']] %>% 
                               pivot_longer(
                                 cols = c("Light1","Light2","Moist1","Moist2","pH1","pH2","Nitrogen1","Nitrogen2"),
                                 names_to = NULL,
                                 values_to = "original",
                                 values_drop_na = FALSE
                               ) %>%
                               pull(original)
  )

head(res.wetland[,70:76])

#### scaled value maps ####
# keep wide format and add geometry again
res.wetland <- results.wetland[['2-sided']]
st_geometry(res.wetland) <- st_geometry(ANO.wetland)

nor <- readRDS('P:/41201785_okologisk_tilstand_2022_2023/data/rds/norway_outline.RDS')%>%
  st_as_sf()

## scaled value maps
# Moist1 (lower indicator)
tm_shape(nor) +
  tm_fill('GID_0', labels="", title="") + #tm_borders() +
  tm_shape(res.wetland) +
  tm_dots('Moist1',midpoint=NA, palette=tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE), scale=1, legend.show = FALSE) + # 
  tm_layout(main.title = "Moisture index (lower), wetland",legend.position = c("right", "bottom"), main.title.size=1.2) + 
  tm_add_legend(type = "fill", 
                col = c(tmaptools::get_brewer_pal("YlOrRd", 5, plot = FALSE),'grey'),
                labels = c("0.4", "...", "...", 
                           "...", "1", "NA"),
                title = "index values")

# Moist2 (upper indicator)
tm_shape(nor) +
  tm_fill('GID_0', labels="", title="") + #tm_borders() +
  tm_shape(res.wetland) +
  tm_dots('Moist2',midpoint=NA, palette=tmaptools::get_brewer_pal("YlOrRd", 7, plot = FALSE), scale=1, legend.show = FALSE) + # 
  tm_layout(main.title = "Moisture index (upper), wetland",legend.position = c("right", "bottom"), main.title.size=1.2) + 
  tm_add_legend(type = "fill", 
                col = c(tmaptools::get_brewer_pal("YlOrRd", 5, plot = FALSE),'grey'),
                labels = c("0.4", "...", "...", 
                           "...", "1", "NA"),
                title = "index values")
